### 2020-12-29 ch/2-related-work.tex
Work on causality section

This can be seen as a form of transfer learning and notably does not exhibit a fairness-accuracy trade-off.

@@

While the algorithms by \citet{feldman2015certifying} are manually constructed and explicit,
\citet{louizos2015variational} use an approach
that falls more in the area of end-to-end learning where fewer hand-crafted algorithms are used.
The method is based on deep variational autoencoders (VAE).
There is an encoder and a decoder that are both modeled as deep neural networks.
The encoder produces the distribution of the latent (fair) representation \(z\)
from the original features \(x\) and the sensitive attribute \(s\).
The decoder recovers the distribution of \(x\) from \(z\) and \(s\).
By choosing a factorized prior \(P(s)P(z)\), a separation between \(s\) and \(z\) is encouraged.

@@

\citet{woodworth2017learning} also consider Equalized Odds as a fairness
definition but give strong criticism of the post-processing approach of Hardt et al.~They
present examples where finding the optimal (but discriminatory) predictor in a particular
hypothesis class and then correcting it post hoc performs poorly. The proofs are based on
the PAC learning framework. Certain training distributions allow the training of good fair
classifiers from scratch but when doing post-processing on a Bayes-optimal predictor for
this distribution, the result can be a very bad predictor that gives essentially random
predictions. The authors explicitly construct one such distribution: it corresponds to the
case where the sensitive attribute \(s\) is more predictive of the class label \(y\) than the
features \(x\). This is arguably an unusual case.

The proposed method is then as follows. First, a classifier is trained that minimizes a
loss function with Equalized Odds as a constraint. The constraint specifies that the
absolute differences in true positive rate and false positive rate between the two groups
are below a certain threshold \(\alpha\). In a second step, post-processing similar to Hardt
et al.~is performed with a holdout training set. This ensures that the predictor does not
have the same weaknesses as the proposal by Hardt et al.~in the given example situations.
However, this more thorough approach is computationally unfeasible.
For a feasible approximation,
\citet{woodworth2017learning} refer to the idea by \citet{zafar2017fairnesstreatment}
of using the covariance as a proxy for statistical
dependency. They refer to this criterion as \emph{Equalized Correlations}. Based on this, they
investigate linear predictors and find that for these, post-processing can achieve
Equalized Correlations as well as training a fair predictor from scratch. Therefore, their
criticism is only valid for non-linear predictors or linear predictors that try to enforce
the actual Equalized Odds. There are no experiments in the paper.

Another approach that aims to define a better proxy for the fairness criteria is \citet{quadrianto2017recycling}.
In that paper, \emph{distribution matching} is used to evaluate fairness.
More specifically, like \citet{louizos2015variational}, Maximum Mean Discrepancy (MMD) is used to check if the
distributions \(P(\hat{y}|s=0)\) and \(P(\hat{y}|s=1)\) are close, to ensure Demographic
Parity. This can also be done for Equality of Opportunity and Equalized Odds with the
appropriate distributions.
\citet{quadrianto2017recycling} show that the approach by \citet{zafar2017fairnessconstraints} is a special case of this.
Furthermore, in contrast to Zafar et al.'s approach the
sensitive attribute is explicitly made available to the classifier during training, but
not during decision time (after training is finished). This is done via \emph{privileged
learning}. Privileged learning is a variant of SVMs where certain features are only
available during training but not during decision time. This is usually used in the
context where some features are costly or impractical to gather. An example is using (only
during training) thermal images in addition to normal images for face recognition. The
additional information is intended to enable the classifier to learn the desired task
quicker or better.

Combining privileged learning and the MMD fairness criterion, the task becomes a
multi-objective optimization for an SVM. The MMD criterion is generally non-convex but
there are several ways to deal with this. Quadrianto and Sharmanska find the best
algorithm (low discrimination, high accuracy) with an evolutionary multi-objective
optimization. The evolutionary algorithm produces a number of variants of the classifier
with different trade-offs between classification error and discrimination. As Zafar et al.
before, the Adult/Census Income dataset is used to test Demographic Parity and the
ProPublica/COMPAS dataset is used for Equality of Opportunity and Equalized Odds. Zafar et
al.'s work was used as a baseline but the implementation failed on the Adult/Census Income
dataset. On the ProPublica/COMPAS dataset, the proposed algorithm performed within error
just as well as Zafar et al.'s algorithm.

With the recent emergence of adversarial neural networks, interest was renewed in creating
fair representations. \citet{madras2018learning} present one such recent
approach.
Madras et al.  convert the statistical criteria Demographic Parity, Equality of Opportunity and Equalized
Odds to adversarial objective functions. The general idea is that the ``adversary'' tries
to gain discriminatory information from the fair representation. More specifically, there
are 4 neural networks: the encoder \(f(x, s)\) that encodes the unfair representation \(x\)
into a fair representation \(z\), the classifier \(g(z)\) that tries to predict \(y\) from \(z\),
the decoder \(k(z, s)\) that tries to decode \(z\) back into \(x\) while knowing about the
sensitive attribute \(s\) and finally the adversary \(h(z)\) that tries to predict \(s\) from
\(z\). The classifier ensures that \(z\) contains enough information to predict \(y\). The
decoder ensures that the fair representation contains enough information from \(x\) and can
be used by other classifiers as well.

The key contribution by Madras et al.~is the objective functions for the adversary. In the
case of Demographic Parity, we take -- for each sensitive group separately -- the average
absolute difference between what the adversary predicted \(h(z)\) and the true sensitive
attribute \(s\). After computing these two separate averages, the averages are added and get
a negative sign to form the objective function. For Equalized Odds, the average absolute
difference is calculated on each sensitive group-label combination \((s, y)\) for \(s \in \{0, 1\}\) and \(y \in \{0, 1\}\) separately. Computing these separate averages is referred
to as \emph{group normalizing} in the paper. The authors present a proof for some upper bounds
on unfairness by the optimal adversary when these objective functions are used. An
alternative objective function could be based on cross-entropy, but Madras et al.~dismiss
this on the grounds that there are situations in which it leads to wrong results. However,
in the experiments they show that the proposed objective and one objective based on
cross-entropy lead to very similar results.

Based on the weights of the different parts of the loss function (adversarial loss,
classification loss, decoder loss, etc.), different trade-offs can be achieved. There are
experiments on the Adult/Census Income dataset and the Heritage Health dataset, but they
are not compared to other works.

\section{Recent developments in Fairness}\label{recent-developments-in-fairness}
At this point, fairness in machine learning has come a long way.
There are several working implementations for various fairness criteria
that give reasonable performance.
Nevertheless, there are still many unexplored areas.

There is, for example, a lot of other recent work on fair representation
with adversarial neural networks.
These approaches vary in the structure of the network and the objective function.
These works are, however, all right now only in pre-print.
There is \citet{edwards2015censoring}, \citet{beutel2017data}
and \citet{zhang2018mitigating} for learning fair representations with adversaries.
\citet{calmon2017optimized} improve on the work by \citet{zemel2013learning}.

\citet{kilbertus2018blind} recently proposed a mechanism
where the sensitive attributes are encrypted to avoid being used for discrimination.
\citet{bechavod2017penalizing} use (as some others before) a regularization approach to train fair classifiers.

Finally, \citet{everitt2017reinforcement} investigate
a problem in reinforcement learning that is similar to bias in classification.
In the previously discussed papers,
the bias originates from the training inputs or the training labels.
In the work by \citet{everitt2017reinforcement},
they consider the case of reinforcement learning (RL) with a corrupt reward signal.
In this case, the RL algorithm might know that the reward is not correct,
but does not know in what specific way it is not correct.
The paper discusses some strategies to deal with this,
but there are very many open questions.

>>>>

% This can be seen as a form of transfer learning and notably does not exhibit a fairness-accuracy trade-off.

@@

While the algorithms by \citet{feldman2015certifying} are manually constructed and explicit,
\citet{louizos2016variational} use an approach
that falls more in the area of end-to-end learning where fewer hand-crafted algorithms are used.
The method is based on deep variational autoencoders (VAE).
There is an encoder and a decoder that are both modeled as deep neural networks.
The encoder produces the distribution of the latent (fair) representation \(z\)
from the original features \(x\) and the sensitive attribute \(s\).
The decoder recovers the distribution of \(x\) from \(z\) and \(s\).
By choosing a factorized prior \(P(s)P(z)\), a separation between \(s\) and \(z\) is encouraged.

@@

\citet{woodworth2017learning} also consider Equalized Odds as a fairness definition
but give strong criticism for the post-processing approach from \citet{hardt2016equality}.
They present examples where finding the optimal (but discriminatory) predictor in a particular hypothesis class
and then correcting it post hoc performs poorly.
The proofs are based on the PAC learning framework.
Certain training distributions allow the training of good fair classifiers from scratch
but when doing post-processing on a Bayes-optimal predictor for this distribution,
the result can be a very bad predictor that gives essentially random predictions.
The authors explicitly construct one such distribution:
it corresponds to the case where the sensitive attribute \(s\) is more predictive of the class label \(y\)
than the features \(x\), which is arguably an unusual case.

The proposed method is then as follows.
First, a classifier is trained that minimizes a loss function with Equalized Odds as a constraint.
The constraint specifies that the absolute differences in true positive rate and false positive rate
between the two groups are below a certain threshold \(\alpha\).
In a second step, post-processing similar to \citet{hardt2016equality} is performed with a holdout training set.
This ensures that the predictor does not have the same weaknesses
as the proposal from \citet{hardt2016equality} in the given example situations.
However, this more thorough approach is computationally unfeasible.
For a feasible approximation, \citet{woodworth2017learning} refer to the idea from \citet{zafar2017fairnesstreatment}
of using the covariance as a proxy for statistical dependency.
They refer to this criterion as \emph{Equalized Correlations}.
Based on this, they investigate linear predictors and find that for these,
post-processing can achieve Equalized Correlations as well as training a fair predictor from scratch.
Therefore, their criticism is valid only for non-linear predictors
or linear predictors that try to enforce the actual Equalized Odds.
There are no experiments in the paper.

Another approach that aims to define a better proxy for the fairness criteria is \citet{quadrianto2017recycling}.
In that paper, \emph{distribution matching} is used to evaluate fairness.
More specifically, like \citet{louizos2016variational}, Maximum Mean Discrepancy (MMD) is used to check
if the distributions \(P(\hat{y}|s=0)\) and \(P(\hat{y}|s=1)\) are close, to ensure Demographic Parity.
This can also be done for Equality of Opportunity and Equalized Odds with the appropriate distributions.
\citet{quadrianto2017recycling} show that the approach by \citet{zafar2017fairnessconstraints} is a special case of this.
Furthermore, in contrast to that approach,
the sensitive attribute is explicitly made available to the classifier during training,
but not during decision time (after training is finished).
This is done via \emph{privileged learning}.
Privileged learning is a variant of \acp{SVM}
where certain features are only available during training but not during decision time.
This is usually used in the context where some features are costly or impractical to gather.
An example is using (only during training) thermal images in addition to normal images for face recognition.
The additional information is intended to enable the classifier to learn the desired task quicker or better.

Combining privileged learning and the MMD fairness criterion,
the task becomes a multi-objective optimization for an SVM.
The MMD criterion is generally non-convex but there are several ways to deal with this.
The authors find the best algorithm (low discrimination, high accuracy)
with an evolutionary multi-objective optimization.
The evolutionary algorithm produces a number of variants of the classifier
with different trade-offs between classification error and discrimination.
As in \citet{zafar2017fairnesstreatment} before,
the Adult/Census Income dataset is used to test Demographic Parity
and the ProPublica/COMPAS dataset is used for Equality of Opportunity and Equalized Odds.
\citet{zafar2017fairnesstreatment} was used as a baseline
but the implementation failed on the Adult/Census Income dataset.
On the ProPublica/COMPAS dataset,
the proposed algorithm performed within error just as well as the fair baseline algorithm.

With the emergence of adversarial neural networks, interest was renewed in creating fair representations.
\citet{ganin2016domain} can be seen as a direct precursor:
they propose adversarial learning for domain adaptation.
Their method involves learning a shared representation that is invariant to the different domains.
If we consider demographic groups as different domains, then this is essentially fair representation learning.

This is the premise of \citet{edwards2016censoring}.
Four neural networks are trained in an adversarial setting:
the encoder \(f(x)\) which encodes the unfair representation \(x\) into a fair representation \(z\),
the classifier \(g(z)\) which tries to predict \(y\) from \(z\),
the decoder \(k(z)\) which tries to reconstruct \(x\) from \(z\),
and finally the adversary \(h(z)\) which tries to predict \(s\) from \(z\).
The classifier \(g\) ensures that \(z\) contains enough information to predict \(y\).
The decoder \(k\) ensures that the fair representation contains enough information from \(x\)
and can be used by other classifiers as well.
The sign of the adversary's loss is inverted for the gradient of the encoder \(f\),
so that encoder tries to make the adversary's predictions worse.
Encoder and adversary should converge to a point
where the adversary can only predict the correct \(s\) from \(z\) at chance level.
Assuming that the adversary is powerful enough to detect any trace of \(s\) in \(z\),
this means that no information about the sensitive attribute \(s\) remains in the fair representation \(z\).
Note that there is a tension here between requiring \(x\) to be reconstructible from \(z\)
and requiring \(z\) to contain no information about \(s\), since \(x\) \emph{does} contain information about \(s\).

\citet{beutel2017data} explicitly connect the method presented in \citet{edwards2016censoring}
to various fairness definitions.

\citet{zhang2018mitigating} propose an architecture similar to \citet{edwards2016censoring},
but with the difference that no decoder (and no reconstruction loss) is used
and that the adversary tries to predict \(s\) from the final output of the classifier
instead of an intermediate representation.
The goal is not to learn a fair representation, but a fair classifier.

\citet{madras2018learning} is another, similar approach.
One difference to \citet{edwards2016censoring} is that several fairness criteria are considered:
Demographic Parity, Equality of Opportunity and Equalized Odds are all converted to adversarial objective functions.
Another difference is that the decoder \(k\) has access to \(s\) for the decoding: \(k: (Z, S)\to X\).
This ensures that there is no tension between reconstructing \(x\) and removing \(s\) from \(z\).
However, the key contribution is the objective functions for the adversary.
In the case of Demographic Parity, we take -- for each sensitive group separately --
the average absolute difference between what the adversary predicted \(h(z)\) and the true sensitive attribute \(s\).
After computing these two separate averages,
the averages are added and get a negative sign to form the objective function.
For Equalized Odds,
the average absolute difference is calculated on each sensitive group-label combination
\((s, y)\) for \(s \in \{0, 1\}\) and \(y \in \{0, 1\}\) separately.
Computing these separate averages is referred to as \emph{group normalizing} in the paper.
The authors present a proof for upper bounds on unfairness by the optimal adversary
when these objective functions are used.
An alternative objective function could be based on cross-entropy -- as used in \citet{edwards2016censoring} --
but the authors dismiss this on the grounds that there are situations in which it leads to wrong results.
However, in the experiments they show
that the proposed objective and one objective based on cross-entropy lead to very similar results.

Based on the weights of the different parts of the loss function
(adversarial loss, classification loss, decoder loss, etc.),
different trade-offs can be achieved.

\section{Fairness via causal reasoning}%
\label{sec:fairness-definition-based-on-causal-reasoning}
In this section, I will briefly discuss a view that also falls under the definition-centric view:
in it, the goal is to make decisions that conform to causality-based notions of fairness.
Methods in this category assume that the causal structure of the task is known,
which can then be used to identify unfair path ways for making decisions.
One motivating example is this \citep{dedeo2014wrong}:
In college admissions, we might not want to discriminate against prospective students from poorer backgrounds.
If the goal is to admit those students that have a chance of graduating,
then we could employ a machine learning algorithm to predict graduation rate.
One attribute that might turn out to be predictive of graduation is physical fitness.
If this attribute's causal influence is through,
for example, signaling a character trait such as grit,
then this is an admissible attribute.
However, if on the other hand physical fitness is caused by access to expensive gyms,
then it is a signal for high socioeconomic status which we do not want to use as a criterion.
So, depending on what the causal mechanism is,
physical fitness could be a discriminatory or a non-discriminatory feature.

% =====================================
%          copied in - start
% =====================================

In this spirit, \citet{kilbertus2017avoiding} use causal graphs to develop new fairness criteria.
In their work, the prediction output \(\hat{y}\) is a part of the causal graph that models the data.
\citet{hardt2016equality} had already defined
Demographic Parity and Equalized Odds (with Equality of Opportunity as a special case)
as statistical independencies;
Demographic Parity: \(s \bot \hat{y}\), Equalized Odds: \(s \bot \hat{y} | y\).
In causal graphs, the first is equivalent to no existing path between \(s\) and \(\hat{y}\),
and the second independence allows paths to exist between \(s\) and \(\hat{y}\)
as long as all paths are all blocked by \(y\).
Instead of considering paths that start from \(s\) directly,
the authors argue that the observed proxies for \(s\) are more important.
A proxy for \(s\) is a clearly defined observable quantity
that is significantly correlated with \(s\).
According to the authors,
the sensitive attribute in its pure conceptual form may influence the prediction directly,
but any observable proxy of it may not.
The reason for this seems to be that if all influence of \(s\) on \(\hat{y}\) was removed,
only very few features would remain to make a decision
because usually nearly all features are influenced by \(s\) somehow.

The concept of removing causal influence
can be efficiently expressed with \emph{interventions} on the causal graph
(level 2 according to Pearl).
An intervention on a variable \(v\) cuts off all parents of \(v\)
and sets \(v\) to a specific value, say \(v'\).
This is written as \(do(v=v')\).
So, for a given proxy \(p\) (\(p \in \{0, 1\}\)),
predictions \(\hat{y}\) exhibit no \emph{proxy discrimination} if
\begin{align}
  \label{eq:proxy-discrimination}
  P(\hat{y}|do(p=0)) = P(\hat{y}|do(p=1))~.
\end{align}
If this is fulfilled, there is no influence from \(s\) that is mediated through \(p\).
There might still be direct influence of \(s\) on \(\hat{y}\).
The fact that \(do\)-interventions are used in the definition means
that a causal model is required to evaluate this criterion; observational data does not suffice.

The causal model that is used in the paper is a structural equation model.
Each edge in the graphical model corresponds to an equation.
There is a node for each input feature, each sensitive attribute (usually only one),
each proxy and the output of the classifier.
Additionally, there is an unobserved error variable for each input feature (and each proxy).
The equation that defines the classification result in terms of the inputs is parameterized.
This equation represents the classifier \(\hat{y} = g_{\theta}(x,p,s)\).
In the paper, only linear functions are used in this equation;
meaning that the classifier is linear.

The following is then the procedure to construct a fair classifier:

\begin{enumerate}
  \item
    Do an intervention with \(do(p=p')\). Change the structural equations accordingly.
  \item
    Construct one big equation for the classifier output \(g_{\theta}(x,p',s) = ...\)
    by substituting all variables according to the graph up to the root nodes,
    taking into account the \(do\)-intervention.
  \item
    Find the set of values of \(\theta\) that fulfills the fairness criterion:
    \(g_{\theta}(x,p'=0,s) = g_{\theta}(x,p'=1,s)\).
    If there is no such value for \(\theta\), the procedure does not work.
  \item
    Within the allowed range of values for \(\theta\),
    find the one that minimizes the classification loss.
\end{enumerate}

The simplest way to achieve step 3 is usually to choose a \(\theta\)
which makes all contributions from descendants of \(p\) vanish,
but the procedure also allows for more sophisticated ways to remove the influence of \(p\).
(Descendants of \(p\) are all variables that can be reached from \(p\) via directed paths.)

% \includegraphics{\{\{ site.url \}\}/blog/assets/fairness-literature/all-connected.svg}
% \includegraphics{\{\{ site.url \}\}/blog/assets/fairness-literature/p-separated.svg}

The resulting classifier is fair by construction.
However, note that this procedure does not lead to true individual fairness,
because the influence of \(p\) is only removed at the population level
(except in the case where all descendants of \(p\) are completely removed).
Individual decisions can still be unfair as long as it balances out overall.

The authors provide a number of properties of their method with corresponding proofs.
For example, if the influence of \(p\) on \(x\) is additive and linear,
any predictor of the form \[g(x - \mathbb{E}[x|do(p=p')])\] has no proxy discrimination.
There are no experiments presented in the paper.

The main weakness of the presented approach are
the need for a causal model that can not be inferred from just the training data
and the difficulty of finding a permissible range of values for \(\theta\).
If the simple way is taken for choosing the range (removing all descendants of \(p\)),
then there might not be enough input features left for learning.
Another weakness is that no true individual fairness is achieved.

True individual causal fairness is, however, the goal of \citet{kusner2017counterfactual}.
In their paper, a new fairness criterion which they call ``counterfactual fairness''
is defined making use of causal reasoning.
For the definition, they assume that we have a causal model
that can be used to compute counterfactuals (corresponding to level 3 in Pearl's hierarchy).
This is in contrast to \citet{kilbertus2017avoiding} who relied on interventions (level 2).
The causal model depends on unobserved background variables \(U\),
non-sensitive features \(x\) and a sensitive attribute \(s\).
A predictor \(\hat{y}\) is then \emph{counterfactually fair}
if the following equality of counterfactual probabilities holds:
\begin{align}
  \label{eq:counterfactual-fairness}
  P(\hat{y}_{s=i}(U) = 1|x, s=i)=P(\hat{y}_{s=j}(U) = 1|x, s=i)
\end{align}
where \(i, j \in \{0, 1\}\) and \(i \neq j\).

The first counterfactual probability is actually not a real counterfactual:
it is just the probability of observing \(\hat{y} = 1\)
given the background variables \(U\), the features \(x\) and \(s=i\).
The second counterfactual probability is the probability of observing \(\hat{y} = 1\)
given \(U\) and \(x\) and given that we had \(s=j\) instead of the actual value \(s=i\).
In other words, the criterion demands this:
the prediction would have been the same in the counterfactual world
where sensitive attribute is \(j\) instead of \(i\).
In the counterfactual world, everything that is not causally dependent on \(s\) is held constant.

The difficult problem that Pearl solved with his theory of counterfactuals
is how to define the counterfactual world where the value of \(s\) is \(j\) instead of \(i\).
One approach that philosophers took in trying to define it
is looking for the ``closest world'' to ours where \(s=j\), according to some measure of distance.
Pearl's counterfactuals do not work this way, but it is a useful intuition.
Making sure predictions are the same, for the closest world with \(s=j\),
is arguably what \citet{dwork2012fairness} (see above) did
with their idea of treating similar people similarly (regardless of \(s\)).
In this way, counterfactual fairness is similar to the individual fairness
that \citet{dwork2012fairness} defined.
However, instead of using a similarity metric,
a causal model is needed for counterfactual fairness.
The two criteria share some strengths and weaknesses.
A strength is that they are both fair on the individual level,
which is not the case for Demographic Parity and Equality of Opportunity.
In Demographic Parity and Equality of Opportunity,
we can achieve fairness by doing ``negative'' discrimination on some individuals
and ``positive'' discrimination on others, as long as those two effects cancel out.
This is because those criteria are statistical criteria.
However, in turn this implies that individual fairness and counterfactual fairness
could not be used to implement affirmative action (``positive'' discrimination for a group)
and they rely on the features \(x\) being correct.

The authors prove that \(\hat{y}\) is counterfactually fair
if it does not causally depend (as defined by the causal model) on \(s\) or any descendants of \(s\).
This motivates the following algorithm to train a fair predictor.
We assume that we have the causal model and the prior over the background variables \(U\).
We train an classifier \(g_{\theta}(U, x_{\not\succ s})\) that is parameterized by \(\theta\).
\(x_{\not\succ s}\) refers to all features \(x\) that are \emph{not} descendants of \(s\).
Furthermore, we have a loss function that measures how well the class labels \(y\) are predicted.
For training, we optimize the expected loss over \(P(U|x_i, s_i)\)
which is the distribution over the background variables, as defined by the causal model,
given training instances \(x_i\) and \(s_i\).
Note that the number of features that are non-descendants of \(s\) might be very small or even zero.
However, this is not a problem as the classifier \(g_{\theta}(U, x_{\not\succ s})\)
can get the necessary information from \(U\).
The variables in \(U\) are, by definition, non-descendants of \(s\).
Predictions are made via \[\hat{y} = \mathbb{E} [g_{\theta}(U^*, x_{\not\succ s}^*)|x^*, s^*]\].

Constructing the causal model remains the major weakness of this approach.
The authors point to existing literature on the topic and give some additional vague guidelines.
In the experiment section, the algorithm is applied to the Law School Success dataset.
The task is to predict the first year average grade (FYA) in law school
given the GPA before law school and the score on the entrance exam (LSAT).
Also known are race and sex, which are not supposed to be used to make predictions.
Logistic regression is used for the classifier.

The authors construct two causal models.
In both models, race and sex influence all three test scores GPA, LSAT and FYA.
In the first model, there is only one background variable that the authors call \emph{knowledge}.
This variable influences the three test scores but does not influence the race or sex.
The second model is a fully deterministic model
where we have three background variables which play the role of additive error terms.
The three error terms may be correlated.
If the underlying causal model is taken as true,
the trained fair classifiers are automatically completely counterfactually fair.
The second model achieves a slightly lower prediction error than the first model
(RMSE of 0.918 vs 0.929).
Additionally, two (unfair) baseline models were trained:
one trained on only \(x\) and the other trained on \(x\) and \(s\).
Those baseline models indeed turn out to be not counterfactually fair.

The major advantages of the approach are:
It enforces true individual fairness and it leaves open the choice of classifier.
This alone makes it one of the best approaches to date.

% =====================================
%          copied in - end
% =====================================

% \section{Recent developments in Fairness}\label{recent-developments-in-fairness}
% At this point, fairness in machine learning has come a long way.
% There are several working implementations for various fairness criteria
% that give reasonable performance.
% Nevertheless, there are still many unexplored areas.

% There is, for example, a lot of other recent work on fair representation
% with adversarial neural networks.
% These approaches vary in the structure of the network and the objective function.
% These works are, however, all right now only in pre-print.
% There is \citet{edwards2016censoring}, \citet{beutel2017data}
% and \citet{zhang2018mitigating} for learning fair representations with adversaries.
% \citet{calmon2017optimized} improve on the work by \citet{zemel2013learning}.

% \citet{kilbertus2018blind} recently proposed a mechanism
% where the sensitive attributes are encrypted to avoid being used for discrimination.
% \citet{bechavod2017penalizing} use (as some others before) a regularization approach to train fair classifiers.

% Finally, \citet{everitt2017reinforcement} investigate
% a problem in reinforcement learning that is similar to bias in classification.
% In the previously discussed papers,
% the bias originates from the training inputs or the training labels.
% In the work by \citet{everitt2017reinforcement},
% they consider the case of reinforcement learning (RL) with a corrupt reward signal.
% In this case, the RL algorithm might know that the reward is not correct,
% but does not know in what specific way it is not correct.
% The paper discusses some strategies to deal with this,
% but there are very many open questions.
